---
title: "Kaggle_description"
author: "Hongxu Zhu"
always_allow_html: yes
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---

```{r alternative-preambles,echo=FALSE,include=FALSE,eval=FALSE}
## if using word replace output with
output:
  word_document:
    toc: yes

## if using pdf replace output with
output:
  pdf_document:
    toc: yes
header-includes:
  - \usepackage{xcolor}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## colorizes text when using html or pdf output
## just spits out text x for word
colorize <- function(x, color) {
  if(color=="todo"){
    color <- "red"
    x <- paste0("TODO: ",x)
  }
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}

todo <- function(x){
  paste0("**",colorize(x,"todo"),"**")
}

#library(kableExtra)
#library(cmapR)
#library(edgeR)
#library(arrow)
```

## Data Description

Let $y^{ca}_{ijkl}\in R^p$ denote a single cell expression sample. Here, $(c,a)$ represents a unique context action pair, $i\in\{0,1,2\}$ is the indicator of donor, $j\in\{0,1,...,5\}$ indicates the plate, $k\in\{A,B,...,H\}$ indicates the row on the plate, while $l$ indicates the single cell sample. We may refer the subscriptions $\{ijk\}$ as the indicator of "location" for a $(c,a)$ pair. Notice that, the location indicators are not independent. The plate indicator $j$ is associated with the donor indicator $i$. In this data set, $j|i=0 \in \{0,2\}$, $j|i=1\in\{3,5\}$ and $j|i=2\in\{1,4\}$. For every location defined by $\{ijk\}$, experiments were applied on every cell type.  

### Control experiments

In this experiment, the first 3 columns for every plate are assigned for control experiments (DMSO, Dabrafenib and Belinostat). Thus, for a given control action for donor $i\in\{0,1,2\}$, there will be 16 possible locations for this control action. For example, let's denote DMSO as $a_0$, then, the samples for $\{ca_0\}$ for donor 0 could be written as $y^{ca_0}_{ijkl}$ where $i=0, j\in\{0,2\},k\in\{A,B,...,H\}$. The total number of possible location would be $2\times 8=16$.

### Perturbation experiments

For every donor $i\in\{0,1,2\}$, each perturbation action is applied in one of the remaining wells (wells that not for control experiments) in the two plates assigned for the donor. Thus, for a given perturbation action $a$, when $i$ is determined, there will be only one possible combination of $j$ and $k$, which results in 3 possible locations for this action. 

```{r echo=FALSE, out.width = "80%", fig.align = "center"}
knitr::include_graphics("kaggle_desc1.pdf")
```

## Pseudo-bulking

Pseudo-bulking was performed using sum aggregation for each cell type under a specific group defined by the location indicator $\{ijk\}$. Denote The pseudo-bulked level expression as $\tilde{y}^{ca}_{ijk}$. This procedure could be written as

$$\tilde{y}^{ca}_{ijk}=\sum_{l=1}^{n^{ca}_{ijk}}y^{ca}_{ijkl}$$ 

where $n^{ca}_{ijk}$ is the total number of single cell samples in group $\{ijk\}$ for a context-action pair $(c,a)$. After pseudo-bulking, we get 48 pseudo-bulked level samples for each control pair and 3 pseudo-bulked level samples for each perturbed pair. However, due to certain practical issues (e.g. no results for certain cell types in a well), we end up have less pseudo-bulked level samples for some $(c,a)$ pair.

## Data visibility and pre-processing

We cannot observe the experiment results for all the $(c,a)$ pairs. These experiment results were masked by the organizer. The detail of this could be viewed in the heatmap. For the remaining observed experiment results, we conduct the following cleaning process:

1. Separate the dataset into control and perturbed

2. No further processing on the control data

3. For a perturbed context-action pair $(c,a)$, if the number of total samples is less than 3, remove the pair from the dataset. This process removes about 50 pairs from the dataset.

```{r echo=FALSE, out.width = "80%", fig.align = "center"}
knitr::include_graphics("kaggle_desc2.pdf")
```

 
## Training and testing split

After the pseudo-bulking and the pre-processing, we have a remaining dataset with 548 perturbed $(c,a)$ pairs. All these pairs have 3 samples that indicated by the location indicator $\{ijk\}$. We split the dataset into training and testing based on the $(c,a)$ pairs: among all the $(c,a)$ pairs in the dataset, we randomly select 100 $(c,a)$ pairs as the testing pairs, the remaining 448 pairs are in the training pairs.

The result of the split returns:

training set: 448 perturbed $(c,a)$ pairs, each pair has 3 samples + control pairs

testing set: 100 perturbed $(c,a)$ pairs

## Fit in silico models

We fit in silico models separately for each donor. (We need to consider the effect of covariates in the model, however, donor is highly correlated with plate). For $m\in\{1,2,...,p\}$, we have

Cell type model: 

$$Y^i_m=\beta_{0m}^i+\beta_{Cm}^iC^i+\beta_{Pm}^iP^i+\beta_{Rm}^iR^i+\varepsilon^i$$

where $i\in\{0,1,2\}$, $C$ is the context (cell type), $P$ is the plate $R$ is the row.

Perturbation model:

$$Y^i_m=\beta_{0m}^i+\beta_{Am}^iA^i+\beta_{Pm}^iP^i+\beta_{Rm}^iR^i+\varepsilon^i$$
$A$ is the action (perturbation).

Two-factor model:

$$Y^i_m=\beta_{0m}^i+\beta_{Cm}^iC^i+\beta_{Am}^iA^i+\beta_{Pm}^iP^i+\beta_{Rm}^iR^i+\varepsilon^i$$
SIA:

1. For each donor, remove plate effect and row effect. 

$$Y^i_m=\beta_{0m}^i+\beta_{Cm}^iC^i+\beta_{Am}^iA^i+\beta_{Pm}^iP^i+\beta_{Rm}^iR^i+\varepsilon^i$$

Remove $\beta_{Pm}^i$ and $\beta_{Rm}^i$.

2. Fit SIA, get prediction.

3. add back batch effects, get final prediction.

## Measure Model performance

Let the predicted expression be $\hat{y}^{ca}_{ijk}$.

1. Compute $R^2$: 

i) $cor(\tilde{y}^{ca}_{ijk},\hat{y}^{ca}_{ijk})^2$ (get $R^2$ for every $i\in\{0,1,2\}$).

ii) $cor(\frac{1}{3}\sum_{i=0}^2\tilde{y}^{ca}_{ijk},\frac{1}{3}\sum_{i=0}^2\hat{y}^{ca}_{ijk})^2$

2. PR-curve and AUC-PR:

Predicted probability (score): $Logfc\times I_{\{p<0.05\}}$ 


## In vitro DEG

In vitro DEG was performed using linear model by Limma. The linear model could be written as: for $m\in\{1,2,...,p\}$

$$Y_m=\beta_{0m}+\beta_{Dm}D+\beta_{Cm}C+\beta_{Am}A+\beta_{Pm}P+\beta_{Rm}R+\varepsilon$$
The p value is obtained by testing the coefficient by contrast.

## In silico DEG

Apply the same linear model using predicted expressions on testing set.









