---
title: "kaggle_analysis"
author: "Hongxu Zhu"
always_allow_html: yes
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: yes
    toc_collapsed: yes
    toc_float: yes
---

```{r alternative-preambles,echo=FALSE,include=FALSE,eval=FALSE}
## if using word replace output with
output:
  word_document:
    toc: yes

## if using pdf replace output with
output:
  pdf_document:
    toc: yes
header-includes:
  - \usepackage{xcolor}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## colorizes text when using html or pdf output
## just spits out text x for word
colorize <- function(x, color) {
  if(color=="todo"){
    color <- "red"
    x <- paste0("TODO: ",x)
  }
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}

todo <- function(x){
  paste0("**",colorize(x,"todo"),"**")
}

library(kableExtra)
library(cmapR)
library(edgeR)
```

## Data input and pre-preocessing

```{r}
load(file="datasets/bulk_counts.Rda")
length(unique(bulk_counts$gene))
length(bulk_counts$bulk_id)
train_meta <- read.csv("datasets/adata_obs_meta.csv")
```

```{r}
library(tidyr)
bc_data <- bulk_counts %>% pivot_wider(names_from = gene, values_from = count)
bc_data[is.na(bc_data)] <- 0
```

```{r}
length(unique(bc_data$bulk_id))
length(unique(paste0(bc_data$cell_type,bc_data$sm_lincs_id)))

```

```{r}
unique(bc_data$sm_lincs_id)
nrow(bc_data[bc_data$sm_lincs_id=="LSM-43181" & bc_data$cell_type=="B cells",1:10])
```

This is a pseudobulked data. Here, pseudobulked means we summed the raw counts for all cells of a given type for each well in the experiment. Following the Kaggle description, the DE gene analysis is performed on pseudobulked data. Here, we fit a limma model including donor, plate, library as covariates.

```{r}
## Figure out lincs id for DMSO, Dabrafenib and Belinostat)
DMSO_id <- unique(train_meta$sm_lincs_id[train_meta$sm_name=="Dimethyl Sulfoxide"])
Dab_id <- unique(train_meta$sm_lincs_id[train_meta$sm_name=="Dabrafenib"])
Bel_id <- unique(train_meta$sm_lincs_id[train_meta$sm_name=="Belinostat"])
```

```{r}
## Control Data sets
DMSO_data <- bc_data[bc_data$sm_lincs_id==DMSO_id,]
Dab_data <- bc_data[bc_data$sm_lincs_id==Dab_id,]
Bel_data <- bc_data[bc_data$sm_lincs_id==Bel_id,]
```

```{r}
test1 <- data.frame(rbind(DMSO_data,Dab_data))
test2 <- data.frame(rbind(DMSO_data,Bel_data))
```

## DE Gene analysis, positive control vs negative control

```{r}
ctype <- unique(test1$cell_type)
temp1 <- test1[test1$cell_type ==ctype[1],]
temp <- test1[test1$cell_type ==ctype[1],-(1:7)]
counts <- t(temp)

d0 <- DGEList(counts)
d0 <- calcNormFactors(d0)
d0

cutoff <- 3
drop <- which(apply(cpm(d0), 1, max) < cutoff)
d <- d0[-drop,] 
dim(d)

condition <- as.character(temp1$sm_lincs_id)
condition[condition==DMSO_id] <- "DMSO"
condition[condition==Dab_id] <- "Dabrafenib"
group <- as.factor(condition)

plate <- temp1$plate_name
donor <- temp1$donor_id
lib <- temp1$library_id


plotMDS(d, col = as.numeric(group))

mm <- model.matrix(~0 + group+plate+donor+lib)

y <- voom(d, mm, plot = T)

fit <- lmFit(y, mm)
head(coef(fit))

contr <- makeContrasts(`groupDabrafenib` - `groupDMSO`, levels = colnames(coef(fit)))
contr
tmp <- contrasts.fit(fit, contr)
tmp <- eBayes(tmp)


top.table <- topTable(tmp, sort.by = "logFC", n = Inf)
head(top.table, 20)

```
```{r, fig.height=5}
library(EnhancedVolcano)
v <- EnhancedVolcano(top.table,
       lab = rownames(d$counts),
       x = 'logFC',
       y = 'adj.P.Val',
       pCutoff=0.01,
       FCcutoff = 1)
print(v)
```

```{r}
ctype <- unique(test2$cell_type)
temp1 <- test2[test2$cell_type ==ctype[1],]
temp <- test2[test2$cell_type ==ctype[1],-(1:7)]
counts <- t(temp)

d0 <- DGEList(counts)
d0 <- calcNormFactors(d0)
d0

cutoff <- 3
drop <- which(apply(cpm(d0), 1, max) < cutoff)
d <- d0[-drop,] 
dim(d)

condition <- as.character(temp1$sm_lincs_id)
condition[condition==DMSO_id] <- "DMSO"
condition[condition==Bel_id] <- "Belinostat"
group <- as.factor(condition)

plate <- temp1$plate_name
donor <- temp1$donor_id
lib <- temp1$library_id


plotMDS(d, col = as.numeric(group))

mm <- model.matrix(~0 + group+plate+donor+lib)

y <- voom(d, mm, plot = T)

fit <- lmFit(y, mm)
head(coef(fit))

contr <- makeContrasts(`groupBelinostat` - `groupDMSO`, levels = colnames(coef(fit)))
contr
tmp <- contrasts.fit(fit, contr)
tmp <- eBayes(tmp)


top.table <- topTable(tmp, sort.by = "logFC", n = Inf)
head(top.table, 20)

```
```{r, fig.height=5}
v <- EnhancedVolcano(top.table,
       lab = rownames(d$counts),
       x = 'logFC',
       y = 'adj.P.Val',
       pCutoff=0.01,
       FCcutoff = 1)
print(v)
```

For convenient, substitute Lincs ID to perturbation name

```{r}
perturbation <- train_meta$sm_name[bc_data$sm_lincs_id]
perturbation[perturbation=="Dimethyl Sulfoxide"] <- "DMSO"
bc_meta <- data.frame(cbind(bc_data[,1:7],perturbation))
bc_expression <- bc_data[,-(1:7)]
```

## Pre-process for model fitting

```{r}
### Normalize count data with limma
### For the whole dataset, consider donor as group factor
count <- t(bc_expression)
dge <- DGEList(count, group = bc_meta$donor_id)
dge <- calcNormFactors(dge)
cutoff <- 3
drop <- which(apply(cpm(dge), 1, max) < cutoff)
dge.filt <- dge[-drop,] 
dim(dge.filt)

count.filt <- dge.filt$counts

## first try manurally log transformation

count.filt.log <- log2(count.filt+1)



#mm <- model.matrix( ~ group, data=dge.filt$samples)
#v <- voom(dge.filt, design = mm, plot = TRUE)

#library(qsmooth)
#qs <- qsmooth(object = count, group_factor = bc_meta$donor_id)
#View(qs@qsmoothData)
```

### Work on log transformed data

```{r}
## create datasets for each donor

log.exp <- data.frame(t(count.filt.log))
log.exp.d0 <- log.exp[bc_data$donor_id=="donor_0",]
log.exp.d1 <- log.exp[bc_data$donor_id=="donor_1",]
log.exp.d2 <- log.exp[bc_data$donor_id=="donor_2",]

meta.d0 <- bc_meta[bc_data$donor_id=="donor_0",]
meta.d1 <- bc_meta[bc_data$donor_id=="donor_1",]
meta.d2 <- bc_meta[bc_data$donor_id=="donor_2",]
```

### Benchmark models for each donor


```{r}
## aggregate expression for each pair
## for donor d0, create data frame for model fitting
cell_type <- meta.d0$cell_type
perturbation <- meta.d0$perturbation
plate <- meta.d0$plate_name
library_id <- meta.d0$library_id
donor <- meta.d0$donor_id
d0.data <- data.frame(cbind(cell_type, perturbation, plate, library_id, donor, log.exp.d0))
cmb <- paste0(cell_type,"+",perturbation)

grp_list <- list()
for (i in 1: length(unique(cmb))){
  grp_list[[i]] <- log.exp.d0[cmb==unique(cmb)[i],]
}

grp_mean_list <- list()
for (i in 1: length(grp_list)){
  if (nrow(grp_list[[i]])==1){
    grp_mean_list[[i]] <- as.numeric(grp_list[[i]])
  }else{
    grp_mean_list[[i]] <- apply(grp_list[[i]],2,mean)
  }
}

d0.aggr <- data.frame()
for (i in 1:length(grp_mean_list)){
  d0.aggr <- data.frame(rbind(d0.aggr,grp_mean_list[[i]]))
}

names(d0.aggr) <- names(log.exp.d0)
cmb.u <- unique(cmb)
cell_type <- sub("\\+.*","",cmb.u)
perturbation <- sub(".*\\+","",cmb.u)

d0.aggr <- data.frame(cbind(cell_type,perturbation,d0.aggr))
d0.aggr.expr <- d0.aggr[,-c(1,2)]
```

```{r}
## Repeat same process wtih d1 and d2
cell_type <- meta.d1$cell_type
perturbation <- meta.d1$perturbation
plate <- meta.d1$plate_name
library_id <- meta.d1$library_id
donor <- meta.d1$donor_id
d1.data <- data.frame(cbind(cell_type, perturbation, plate, library_id, donor, log.exp.d1))
cmb <- paste0(cell_type,"+",perturbation)

grp_list <- list()
for (i in 1: length(unique(cmb))){
  grp_list[[i]] <- log.exp.d1[cmb==unique(cmb)[i],]
}

grp_mean_list <- list()
for (i in 1: length(grp_list)){
  if (nrow(grp_list[[i]])==1){
    grp_mean_list[[i]] <- as.numeric(grp_list[[i]])
  }else{
    grp_mean_list[[i]] <- apply(grp_list[[i]],2,mean)
  }
}

d1.aggr <- data.frame()
for (i in 1:length(grp_mean_list)){
  d1.aggr <- data.frame(rbind(d1.aggr,grp_mean_list[[i]]))
}

names(d1.aggr) <- names(log.exp.d1)
cmb.u <- unique(cmb)
cell_type <- sub("\\+.*","",cmb.u)
perturbation <- sub(".*\\+","",cmb.u)

d1.aggr <- data.frame(cbind(cell_type,perturbation,d1.aggr))
d1.aggr.expr <- d1.aggr[,-c(1,2)]

###############

cell_type <- meta.d2$cell_type
perturbation <- meta.d2$perturbation
plate <- meta.d2$plate_name
library_id <- meta.d2$library_id
donor <- meta.d2$donor_id
d2.data <- data.frame(cbind(cell_type, perturbation, plate, library_id, donor, log.exp.d2))
cmb <- paste0(cell_type,"+",perturbation)

grp_list <- list()
for (i in 1: length(unique(cmb))){
  grp_list[[i]] <- log.exp.d2[cmb==unique(cmb)[i],]
}

grp_mean_list <- list()
for (i in 1: length(grp_list)){
  if (nrow(grp_list[[i]])==1){
    grp_mean_list[[i]] <- as.numeric(grp_list[[i]])
  }else{
    grp_mean_list[[i]] <- apply(grp_list[[i]],2,mean)
  }
}

d2.aggr <- data.frame()
for (i in 1:length(grp_mean_list)){
  d2.aggr <- data.frame(rbind(d2.aggr,grp_mean_list[[i]]))
}

names(d2.aggr) <- names(log.exp.d2)
cmb.u <- unique(cmb)
cell_type <- sub("\\+.*","",cmb.u)
perturbation <- sub(".*\\+","",cmb.u)

d2.aggr <- data.frame(cbind(cell_type,perturbation,d2.aggr))
d2.aggr.expr <- d2.aggr[,-c(1,2)]
```


```{r}
## Cellular model
## get prediction for Foretinib for d0

fit <- lm(as.matrix(d0.aggr.expr) ~ d0.aggr$cell_type, x=TRUE)
ctype <- unique(d0.aggr$cell_type)[-(1:2)]

ct_pred0 <- list()
for (i in 1: length(ctype)){
  ix <- d0.aggr$perturbation=="Foretinib" & d0.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d0.aggr.expr[ix,]) 
  X_train <- as.matrix(d0.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_ct <- obs_test%*%fit.x$coefficients
  names(pred_ct) <- names(log.exp.d0)
  ct_pred0[[i]] <- pred_ct[1,]
}
names(ct_pred0) <- ctype

## get prediction for Foretinib for d1

fit <- lm(as.matrix(d1.aggr.expr) ~ d1.aggr$cell_type, x=TRUE)
ctype <- unique(d1.aggr$cell_type)[-(1:2)]

ct_pred1 <- list()
for (i in 1: length(ctype)){
  ix <- d1.aggr$perturbation=="Foretinib" & d1.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d1.aggr.expr[ix,]) 
  X_train <- as.matrix(d1.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_ct <- obs_test%*%fit.x$coefficients
  names(pred_ct) <- names(log.exp.d1)
  ct_pred1[[i]] <- pred_ct[1,]
}
names(ct_pred1) <- ctype

## get prediction for Foretinib for d2

fit <- lm(as.matrix(d2.aggr.expr) ~ d2.aggr$cell_type, x=TRUE)
ctype <- unique(d2.aggr$cell_type)[-(1:2)]

ct_pred2 <- list()
for (i in 1: length(ctype)){
  ix <- d2.aggr$perturbation=="Foretinib" & d2.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d2.aggr.expr[ix,]) 
  X_train <- as.matrix(d2.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_ct <- obs_test%*%fit.x$coefficients
  names(pred_ct) <- names(log.exp.d2)
  ct_pred2[[i]] <- pred_ct[1,]
}
names(ct_pred2) <- ctype
```

```{r}
## Perturbation model
## get prediction for Foretinib for d0

fit <- lm(as.matrix(d0.aggr.expr) ~ d0.aggr$perturbation, x=TRUE)

pt_pred0 <- list()
for (i in 1: length(ctype)){
  ix <- d0.aggr$perturbation=="Foretinib" & d0.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d0.aggr.expr[ix,]) 
  X_train <- as.matrix(d0.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_pt <- obs_test%*%fit.x$coefficients
  names(pred_ct) <- names(log.exp.d0)
  pt_pred0[[i]] <- pred_pt[1,]
}
names(pt_pred0) <- ctype

## get prediction for Foretinib for d1

fit <- lm(as.matrix(d1.aggr.expr) ~ d1.aggr$perturbation, x=TRUE)

pt_pred1 <- list()
for (i in 1: length(ctype)){
  ix <- d1.aggr$perturbation=="Foretinib" & d1.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d1.aggr.expr[ix,]) 
  X_train <- as.matrix(d1.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_ct <- obs_test%*%fit.x$coefficients
  names(pred_pt) <- names(log.exp.d1)
  pt_pred1[[i]] <- pred_pt[1,]
}
names(pt_pred1) <- ctype

## get prediction for Foretinib for d2

fit <- lm(as.matrix(d2.aggr.expr) ~ d2.aggr$perturbation, x=TRUE)

pt_pred2 <- list()
for (i in 1: length(ctype)){
  ix <- d2.aggr$perturbation=="Foretinib" & d2.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d2.aggr.expr[ix,]) 
  X_train <- as.matrix(d2.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_pt <- obs_test%*%fit.x$coefficients
  names(pred_pt) <- names(log.exp.d2)
  pt_pred2[[i]] <- pred_pt[1,]
}
names(pt_pred2) <- ctype
```

```{r}
## Two-factor model
## get prediction for Foretinib for d0

fit <- lm(as.matrix(d0.aggr.expr) ~ d0.aggr$cell_type + d0.aggr$perturbation, x=TRUE)

tf_pred0 <- list()
for (i in 1: length(ctype)){
  ix <- d0.aggr$perturbation=="Foretinib" & d0.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d0.aggr.expr[ix,]) 
  X_train <- as.matrix(d0.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_tf <- obs_test%*%fit.x$coefficients
  names(pred_tf) <- names(log.exp.d0)
  tf_pred0[[i]] <- pred_tf[1,]
}
names(tf_pred0) <- ctype

## get prediction for Foretinib for d1

fit <- lm(as.matrix(d1.aggr.expr) ~ d1.aggr$cell_type + d1.aggr$perturbation, x=TRUE)

tf_pred1 <- list()
for (i in 1: length(ctype)){
  ix <- d1.aggr$perturbation=="Foretinib" & d1.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d1.aggr.expr[ix,]) 
  X_train <- as.matrix(d1.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_tf <- obs_test%*%fit.x$coefficients
  names(pred_tf) <- names(log.exp.d1)
  tf_pred1[[i]] <- pred_tf[1,]
}
names(tf_pred1) <- ctype

## get prediction for Foretinib for d2

fit <- lm(as.matrix(d2.aggr.expr) ~ d2.aggr$cell_type + d2.aggr$perturbation, x=TRUE)

tf_pred2 <- list()
for (i in 1: length(ctype)){
  ix <- d2.aggr$perturbation=="Foretinib" & d2.aggr$cell_type==ctype[i]
  obs_test <- fit$x[ix,]
  obs_train <- fit$x[!ix,]

  X_test <-as.matrix(d2.aggr.expr[ix,]) 
  X_train <- as.matrix(d2.aggr.expr[!ix,])

  fit.x <- lm.fit(obs_train,X_train)
  pred_tf <- obs_test%*%fit.x$coefficients
  names(pred_tf) <- names(log.exp.d2)
  tf_pred2[[i]] <- pred_tf[1,]
}
names(tf_pred2) <- ctype
```

```{r}
source("sia/SI-A.R")
```

```{r}
library(ggplot2)
obs <- list()
for (i in 1:length(unique(d0.aggr$cell_type))){
  obs[[i]] <- d0.aggr.expr[d0.aggr$perturbation=="Foretinib" & d0.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d0)
R2_ct <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(ct_pred0[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_ct[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(ct_pred0[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_ct[i]))+ theme(plot.tag.position = "topright")
  print(p)
}


```

```{r}
obs <- list()
for (i in 1:length(unique(d1.aggr$cell_type))){
  obs[[i]] <- d1.aggr.expr[d1.aggr$perturbation=="Foretinib" & d1.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d1)
R2_ct <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(ct_pred1[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_ct[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(ct_pred1[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_ct[i]))+ theme(plot.tag.position = "topright")
  print(p)
}
```

```{r}
obs <- list()
for (i in 1:length(unique(d2.aggr$cell_type))){
  obs[[i]] <- d2.aggr.expr[d2.aggr$perturbation=="Foretinib" & d2.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d2)
R2_ct <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(ct_pred2[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_ct[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(ct_pred2[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_ct[i]))+ theme(plot.tag.position = "topright")
  print(p)
}
```

```{r}
obs <- list()
for (i in 1:length(unique(d0.aggr$cell_type))){
  obs[[i]] <- d0.aggr.expr[d0.aggr$perturbation=="Foretinib" & d0.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d0)
R2_pt <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(pt_pred0[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_pt[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(pt_pred0[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_pt[i]))+ theme(plot.tag.position = "topright")
  print(p)
}

```

```{r}
obs <- list()
for (i in 1:length(unique(d1.aggr$cell_type))){
  obs[[i]] <- d1.aggr.expr[d1.aggr$perturbation=="Foretinib" & d1.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d1)
R2_pt <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(pt_pred1[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_pt[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(pt_pred1[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_pt[i]))+ theme(plot.tag.position = "topright")
  print(p)
}
```

```{r}
obs <- list()
for (i in 1:length(unique(d2.aggr$cell_type))){
  obs[[i]] <- d2.aggr.expr[d2.aggr$perturbation=="Foretinib" & d2.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d2)
R2_pt <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(pt_pred2[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_pt[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(pt_pred2[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_pt[i]))+ theme(plot.tag.position = "topright")
  print(p)
}
```

```{r}
obs <- list()
for (i in 1:length(unique(d0.aggr$cell_type))){
  obs[[i]] <- d0.aggr.expr[d0.aggr$perturbation=="Foretinib" & d0.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d0)
R2_tf <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(tf_pred0[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_tf[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(tf_pred0[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_tf[i]))+ theme(plot.tag.position = "topright")
  print(p)
}

```

```{r}
obs <- list()
for (i in 1:length(unique(d1.aggr$cell_type))){
  obs[[i]] <- d1.aggr.expr[d1.aggr$perturbation=="Foretinib" & d1.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d1)
R2_tf <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(tf_pred1[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_tf[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(tf_pred1[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_tf[i]))+ theme(plot.tag.position = "topright")
  print(p)
}
```

```{r}
obs <- list()
for (i in 1:length(unique(d2.aggr$cell_type))){
  obs[[i]] <- d2.aggr.expr[d2.aggr$perturbation=="Foretinib" & d2.aggr$cell_type==ctype[i],]
}
genes <- names(log.exp.d2)
R2_tf <- c()
for (i in 1: length(ctype)){
  observation <- as.numeric(obs[[i]])
  prediction <- as.numeric(tf_pred2[[i]])
  dt <- data.frame(genes,observation,prediction)
  names(dt) <- c("genes", "observation", "prediction")
  R2_tf[i] <- round(cor(as.numeric(obs[[i]]),as.numeric(tf_pred2[[i]]))^2,2)
  
 p <- ggplot(dt, aes(x=observation, y=prediction)) + 
 geom_point(shape=2,alpha=0.8, color="blue")+
  geom_smooth(method=lm, na.rm = TRUE,colour="black")+
  xlab("observation") + ylab("Prediction")+ ggtitle(ctype[[i]])+
  labs(tag = paste0("R2=",R2_tf[i]))+ theme(plot.tag.position = "topright")
  print(p)
}
```

### DEG analysis

```{r}
### DEG NK cell type Trametinib vs DMSO
DEG_obs_dmso <- data.frame(t(d0.aggr.expr[d0.aggr$cell_type=="NK cells" & d0.aggr$perturbation=="DMSO",]),t(d1.aggr.expr[d1.aggr$cell_type=="NK cells" & d1.aggr$perturbation=="DMSO",]),t(d2.aggr.expr[d2.aggr$cell_type=="NK cells" & d2.aggr$perturbation=="DMSO",]))

DEG_obs_dmso_full <- data.frame(t(d0.data[d0.data$cell_type=="NK cells" & d0.data$perturbation=="DMSO",]),t(d1.data[d1.data$cell_type=="NK cells" & d1.data$perturbation=="DMSO",]),t(d2.data[d2.data$cell_type=="NK cells" & d2.data$perturbation=="DMSO",]))

DEG_obs_dmso_expr <- DEG_obs_dmso_full[-(1:5),]

#names(DEG_obs_dmso) <- c("donor1", "donor2","donor3")

#################

DEG_obs_For <- data.frame(t(d0.data[d0.data$cell_type=="NK cells" & d0.data$perturbation=="Foretinib",]),t(d1.data[d1.data$cell_type=="NK cells" & d1.data$perturbation=="Foretinib",]),t(d2.data[d2.data$cell_type=="NK cells" & d2.data$perturbation=="Foretinib",]))

#names(DEG_obs_tra) <- c("donor1", "donor2","donor3")

##################

DEG_obs_ct <- data.frame(ct_pred0[[1]],ct_pred1[[1]],ct_pred2[[1]])

#names(DEG_obs_ct) <- c("donor1", "donor2","donor3")

###################

DEG_obs_pt <- data.frame(pt_pred0[[1]],pt_pred1[[1]],pt_pred2[[1]])

names(DEG_obs_pt) <- c("donor1", "donor2","donor3")

###################

DEG_obs_tf <- data.frame(tf_pred0[[1]],tf_pred1[[1]],tf_pred2[[1]])

#names(DEG_obs_tf) <- c("donor1", "donor2","donor3")
```

```{r}

temp <- data.frame(cbind(DEG_obs_dmso_full,DEG_obs_For))
counts <- data.frame(sapply(temp[-(1:5),],as.numeric))
rownames(counts) <- rownames(temp[-(1:5),])
group <- as.factor(t(temp[2,]))
plate <- as.factor(t(temp[3,]))
library_id <- as.factor(t(temp[4,]))
donor <- as.factor(t(temp[5,]))
design <- model.matrix(~group+plate+library_id+donor)
fit <- lmFit(counts,design)
fit1 <- eBayes(fit)
top.table.ob <- topTable(fit1,coef = "groupForetinib",sort.by = "logFC", n = Inf)
deg_ob <- rownames(top.table.ob)[top.table.ob$logFC>=0.3 & top.table.ob$adj.P.Val<0.05]

library(EnhancedVolcano)

v <- EnhancedVolcano(top.table.ob,
       lab = rownames(counts),
       title = "NK",
       x = 'logFC',
       y = 'adj.P.Val',
       pCutoff=0.05,
       FCcutoff = 0.5)
print(v)

################################
temp2 <- data.frame(cbind(DEG_obs_dmso_expr,DEG_obs_ct))
counts <- data.frame(sapply(temp2,as.numeric))
rownames(counts) <- rownames(temp[-(1:5),])
group <- as.factor(t(temp[2,]))
plate <- as.factor(t(temp[3,]))
library_id <- as.factor(t(temp[4,]))
donor <- as.factor(t(temp[5,]))
design <- model.matrix(~group+plate+library_id+donor)
fit <- lmFit(counts,design)
fit1 <- eBayes(fit)
top.table.ct <- topTable(fit1,coef = "groupForetinib",sort.by = "logFC", n = Inf)

deg_ct <- rownames(top.table.ct)[top.table.ct$logFC>=0.3 & top.table.ct$adj.P.Val<0.05]


v <- EnhancedVolcano(top.table.ct,
       lab = rownames(counts),
       title = "NK",
       x = 'logFC',
       y = 'adj.P.Val',
       pCutoff=0.05,
       FCcutoff = 0.5)
print(v)

###################################

temp3 <- data.frame(cbind(DEG_obs_dmso_expr,DEG_obs_tf))
counts <- data.frame(sapply(temp3,as.numeric))
rownames(counts) <- rownames(temp[-(1:5),])
group <- as.factor(t(temp[2,]))
plate <- as.factor(t(temp[3,]))
library_id <- as.factor(t(temp[4,]))
donor <- as.factor(t(temp[5,]))
design <- model.matrix(~group+plate+library_id)
fit <- lmFit(counts,design)
fit1 <- eBayes(fit)
top.table.tf <- topTable(fit1,coef = "groupForetinib",sort.by = "logFC", n = Inf)

v <- EnhancedVolcano(top.table.tf,
       lab = rownames(counts),
       title = "NK",
       x = 'logFC',
       y = 'adj.P.Val',
       pCutoff=0.05,
       FCcutoff = 0.5)
print(v)

deg_tf <- rownames(top.table.tf)[top.table.tf$logFC>=0.3 & top.table.tf$adj.P.Val<0.05]


temp4 <- data.frame(cbind(DEG_obs_dmso_expr,DEG_obs_pt))
counts <- data.frame(sapply(temp4,as.numeric))
rownames(counts) <- rownames(temp[-(1:5),])
group <- as.factor(t(temp[2,]))
plate <- as.factor(t(temp[3,]))
library_id <- as.factor(t(temp[4,]))
donor <- as.factor(t(temp[5,]))
design <- model.matrix(~group+plate+library_id)
fit <- lmFit(counts,design)
fit1 <- eBayes(fit)
top.table.pt <- topTable(fit1,coef = "groupForetinib",sort.by = "logFC", n = Inf)

deg_pt <- rownames(top.table.pt)[top.table.pt$logFC>=0.3 & top.table.pt$adj.P.Val<0.05]


v <- EnhancedVolcano(top.table.pt,
       lab = rownames(counts),
       title = "NK",
       x = 'logFC',
       y = 'adj.P.Val',
       pCutoff=0.05,
       FCcutoff = 0.5)
print(v)
```